---
title: Air Quality Data Available in openair
author: David Carslaw
date: '2019-05-05'
slug: air-quality-data-available-in-openair
summary: A look at the air quality data accessible with openair
categories:
  - R
  - openair
tags: 
  - R
  - openair
image:
  caption: ''
  focal_point: ''
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 1
---

# Access site information

{{% alert note %}} The UK has a surprisingly large amount of air
quality data available. The main national network is called the
[Automatic Urban and Rural
Network](https://uk-air.defra.gov.uk/networks/network-info?view=aurn). There
are however other large networks including the [Scottish Air Quality
Network](http://www.scottishairquality.scot/) and the [Welsh Air
Quality Network](https://airquality.gov.wales/). One of the largest
--- if not the largest urban air quality network is the [London Air
Quality Network](https://www.londonair.org.uk/LondonAir/Default.aspx)
{{% /alert %}}

[openair]{.pkg} includes the function `importMeta` to provide
information on UK air pollution monitoring sites. There are currently
four networks that [openair]{.pkg} has access to:

- The Defra Automatic Urban and Rural Network (AURN)
- The Scottish Air Quality Network (SAQN)
- The Welsh Air Quality Network (WAQN)
- Network(s) operated by King's College London (KCL)

These functions are described in more detail [here](http://davidcarslaw.github.io/openair/reference/index.html).

The `importMeta` function is the first place to look to discover what
sites exist, site type (e.g. traffic, background, rural) and latitude
and longitude.

```{r}
# first load openair
library(openair)

aurn <- importMeta(source = "aurn")

head(aurn)
```

How many of each site type are there?

```{r}
table(aurn$site_type)
```

Sometimes it is necessary to have more information on the sites such
as when they started (or stopped) measuring, the pollutants measured
and the regions in which they exist. Additional site information can
be obtained using the option `all = TRUE`. In the example below, we
will select sites that measure NO~2~ at traffic locations.

```{r warning=FALSE,message=FALSE}
library(tidyverse)
aurn_detailed <- importMeta(source = "aurn", all = TRUE)

no2_sites <- filter(
  aurn_detailed,
  variable == "NO2",
  site_type == "Urban Traffic"
)

nrow(no2_sites)
```

# Plot sites on a map

Since [openair]{.pkg} started there have been huge developments with R
and optional packages. These developments have made it much easier to
manipulate and plot data e.g. with [ggplot2]{.pkg} and the likes of
[dplyr]{.pkg}. There is also now much more focus on interactive
plotting, which is very useful in the context of considering air
pollution sites.

In the example below the unique sites are selected from
`aurn_detailed` because the `site` repeats the number of pollutants
that are measured. Information is also collected for the map popups
and then the map is plotted.

```{r}
library(leaflet)

aurn_unique <- distinct(aurn_detailed, site, .keep_all = TRUE)

# information for map markers
content <- paste(
  paste(
    aurn_unique$site,
    paste("Code:", aurn_unique$code),
    paste("Start:", aurn_unique$date_started),
    paste("End:", aurn_unique$date_ended),
    paste("Site Type:", aurn_unique$site_type),
    sep = "<br/>"
  )
)


# plot map
leaflet(aurn) %>%
  addTiles() %>%
  addMarkers(~ longitude, ~ latitude, popup = content,
             clusterOptions = markerClusterOptions())
```

# Access the data

The information above should help to describe the air quality data
that is easily available through [openair]{.pkg}. Access to the data
is possible through a family of functions that all tend to work in a
similar way. Only two pieces of information are required: the site
code(s) and the year(s) of interest.

So, to import data for the industrial Port Talbot Margam site (close
to a steelworks) with the site code "PT4" for 2015 to 2018, we can:

```{r importPT}
margam <- importWAQN(site = "pt4", year = 2015:2018)
```

This data also includes estimates of wind speed and direction (`ws`
and `wd`) from the WRF model, so we can easily plot the distribution
of concentrations by wind direction. This plot indicates that the
highest PM$_{10}$ concentrations are from the south-west i.e. the
steelworks direction. A better indication of important steelworks
combustion sources can be seen by plotting SO$_2$.

```{r perRose}
percentileRose(margam, 
               pollutant = "pm10", 
               percentile = c(25, 50, 75, 90, 95, 99), 
               cols = "inferno", 
               smooth = TRUE
               )
```

